Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-11-25 Revised-paper/Reply/Reply-Reviewer1.tex

Changes made in response to comments of Reviewer 1: \begin{enumerate}
%%1
\item
\textsf
{Abstract, hightlights, and introduction are written in a way that confuses the reader: the text jumps back and forth between bubbles and drops as well as shear flows and other flow configurations. Then it suddenly moves to lift and deformation rather than breakup. Finally a number of random applications are mentioned without any relation to bubble breakup. The authors should clearly state what their contrbution is from the outset. The reader will not go on until the end to figure that out by himself. The use of shorter paragraphs each addressing a single issue is very helpful.}
\vspace{3 mm}

Response: \\
In the revision, we made many improvements to the grammar and structure so that the reader can follow in a much more efficient manner. \\
\\

%%2
\item
\textsf
{Method: The use of only 2 levels of grid refinement appears odd. For an accurate representation of the flow near the interface a very fine mesh is typically needed. Only 32 cells over the bubble diameter is hardly sufficient ot get accurate results as corroborated by the mediorce agreement of the presented simulation results with the experimental data used for validation. Two more levels of grid refinement are likely needed to resolved the flow near the bubble while farther away some coarsening may be possible. Optimizing the grid is imperative in a situaltion like the considerd one due to the strong deformation of the bubble which high curvature of the interface and the singular nature of the breakup event. Fully capturing the breakup, which represents a topological singularity, is impossible without a subgrid scale model.}
\vspace{3 mm}

Response: \\
		There are two main problems identified from the referee: (a) the grid refinement study does not demonstrate enough convergence and (b) even if we could refine the grid an infinite number of times, our break-up model is insufficient and whatever findings that we have would be completely different than if we were able to model the break-up process with a molecular scale model.  We disagree with the referee; we have added more references to our article in which the uncertainty in our article is smaller than that in the latest existing work. Also we note that existing articles on bubble and drop breakup due to shear do not take into account sub-scale models either and these articles demonstrate reasonable agreement with experiments (as does our submission).  Note: if we were considering the merging of droplets in which bouncing versus merging can occur, then sub-scale drainage can be important to model.  This issue is not relevent for our article though.  We have added this discussion to the introduction.

Finally, we note that our simulations already take about one year.  The bubble deformation case, is a much more challenging problem than the drop case due to the large density ratio.  Anymore refinement would extend the time of simulation to 8 years using current technology.  It is agreed though that speeding up the algorithm in order to run more efficiently would certainly reduce the uncertainty of our results even more.  This is future work.
%Muller-Fischer comparison

### 2024-11-25 Revised-paper/Paper/Bubble-deformation-Rev.tex
working on the paper, forming an argument

Bubble dynamics in shear flow, including breakup, is critically important for various \textcolor{blue} {gas-liquid} scientific and engineering processes.  We refer the reader to the following experimental studies relating to bubble deformation in foaming processes, microfluidic devices, microbubbles in the blood circulation system, materials processing, pharmaceuticals, minerals engineering, bubble column reactor, and cosmetics\cite{ChuFinBouAtaHamPug19,MulTobDreFisWin08,BenRodFauPinFerPerGarMirLim18,DreSai15,EFTEKHARI2021837,WANG2023108105,doi:10.1021/acs.langmuir.1c01814,yoshikawa2010bubble,CanedoETAL}.  In particular, it is the study of bubble deformation as it pertains to high-performance plastics applications that motivate this work.

@@

This article presents computational studies of shear-driven deformation and breakup of a bubble in insoluble viscous liquids.  Studying bubble break-up, in which we focus only on the balance of the wall driving force and bubble surface tension force, via computation rather than (terrestrial) experiments simplifies the process of setting a combination of precise, simple shear flow conditions, low $Ca$ ($Ca=\mu_{m}U/\sigma$) conditions, low-density ratio, low viscosity ratios, and zero buoyancy effects.  The physical properties that distinguish bubble and drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$, where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts ``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'', respectively.  For a bubble in an insoluble, viscous liquid, $\lambda \simeq 0$ and $\eta \simeq 0$, whereas most (terrestrial) studies dealing with a drop in an immiscible viscous liquid take $\lambda =1$ and $\eta \simeq 1$.  

@@

In previous studies on the motion of bubble deformation in a simple shear flow \cite{CanedoETAL,RusMan02, MulTobDreFisWin08}, only findings for bubble deformation under very low $Re$ number conditions ($Re \ll 1$) have been reported.  This is understandable since a low Reynolds number matrix fluid mitigates the effect of gravity on distorting the comparison with the drop case.  In the results reported in \cite{CanedoETAL}, the $x-z$ cross section of the deforming bubble was close to spherical in their experiments.  This is expected since the experiments designed by \cite{CanedoETAL} closely followed the following basic assumptions: (i) ``steady creeping flow with negligible inertial effects,'' (ii) ``Incompressible Newtonian fluids,'' (iii) ``No buoyancy effects,'' (iv) ``No wall effects.''  In this work, we determine, for the first time, the critical Reynolds number ($Re \gg 1$) that leads to bubble breakup.  Additionally, our computational studies reveal characteristics that distinguish a drop's deformation and breakup processes versus those of a bubble.

@@

We remark that there have been a number of articles on the study of lift of slightly deformable bubbles\cite{ErvinANDTryggvason1997,legendre1998lift}.  We reiterate, though, that for bubble deformation and breakup in shear flows, only a few articles exist: \cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}.  These previous studies mainly examined the dynamics (e.g., rotation angle) of bubble deformation in shear flow.  Concerning bubble breakup, Wei et al.~\cite{WeiQiaXu12} presented one numerical result for a bubble breakup process under the condition of $Ca$ (capillary number) = 35. Sharifi et al\cite{AAMM-16-2} presented two results for bubble breakup corresponding to $Ca=7.5$ and $Ca=11.2$.  We point out that all of the previous computational research on bubble deformation (and breakup) under shear driven flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2} use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (e.g. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}, maintains the gas/liquid interface as sharp, enables accurate approximation of the surface tension force,  and by construction the CLSVOF method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer we must make our computational domain (bigger domain aspect ratio!) thereby adversely effecting the condition number even more for carrying our the pressure projection. 

@@

To highlight the mechanisms of bubble deformation and breakup in a shear flow, we juxtapose the bubble results with those of a drop (for terrestrial experiments, when the density ratio of drop to matrix is very close to one!).  We remark that while the study of critical tensile strength parameters for the bubble is sparse, there have been many studies for the simpler drop problem. For completeness, we give a brief overview of previous ``tensile strength'' studies pertaining to drops.

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).

>>>>

Bubble dynamics in shear flow, including breakup, is critically important for various \textcolor{blue} {gas-liquid} scientific and engineering processes.  We refer the reader to the following experimental studies relating to bubble deformation in foaming processes, microfluidic devices, microbubbles in the blood circulation system, materials processing, pharmaceuticals, minerals engineering, bubble column reactor, underwater projectiles, polyethylene devolatilization, microfluidics, food aeration, and cosmetics\cite{ChuFinBouAtaHamPug19,MulTobDreFisWin08,BenRodFauPinFerPerGarMirLim18,DreSai15,EFTEKHARI2021837,WANG2023108105,doi:10.1021/acs.langmuir.1c01814,yoshikawa2010bubble,CanedoETAL,GAO2022103212,WONG2012417,PolyethyleneDevo,FRENSE2024120579,lohse2018bubble,schluter2021small,SANOGO2023112478}.  In particular, it is the study of bubble deformation as it pertains to high-performance plastics applications that motivate this work.

@@

	This article presents computational studies of shear-driven deformation and breakup of a bubble in insoluble viscous liquids.  Studying bubble break-up, in which we focus only on the balance of the wall driving force and bubble surface tension force, via computation rather than experiments simplifies the process of setting a combination of precise, simple shear flow conditions, low $Ca$ ($Ca=\mu_{m}U/\sigma$) conditions, low-density ratio, low viscosity ratios, and zero buoyancy effects.  We remark that the majority of previous experimental studies on shear-driven bubble deformation are restricted to low Reynolds' number and large capillary number\cite{RusMan02,CanedoETAL}.  One reason is that buoyancy effects are minimized in highly viscous liquids.  On the other hand, many applications are characterized by liquids in the high Reynolds' number regime.  The physical properties that distinguish bubble and drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$, where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts ``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'', respectively.  For a bubble in an insoluble, viscous liquid, $\lambda \simeq 0$ and $\eta \simeq 0$, whereas most studies dealing with a drop in an immiscible viscous liquid take $\lambda =1$ and $\eta \simeq 1$.  

@@

In previous studies on the motion of bubble deformation in a simple shear flow \cite{CanedoETAL,RusMan02, MulTobDreFisWin08}, only findings for bubble deformation under very low $Re$ number conditions ($Re \ll 1$) have been reported.  This is understandable since a low Reynolds number matrix fluid mitigates the effect of gravity on distorting the comparison with the drop case.  In the results reported in \cite{CanedoETAL}, the $x-z$ cross section of the deforming bubble was close to spherical in their experiments.  Also in the experiments in \cite{CanedoETAL}, the $x-z$ cross section shows minimal overall bubble rise speed in the (very) viscous fluid in comparison to the rate that the bubble deforms.  This is expected since the experiments designed by \cite{CanedoETAL} closely followed the following basic assumptions: (i) ``steady creeping flow with negligible inertial effects,'' (ii) ``Incompressible Newtonian fluids,'' (iii) ``No buoyancy effects,'' (iv) ``No wall effects.''  In this work, we determine, for the first time, the critical Reynolds number ($Re \gg 1$) that leads to bubble breakup.  Additionally, our computational studies reveal characteristics that distinguish a drop's deformation and breakup processes versus those of a bubble.

@@

We remark that there have been a number of computational articles on the study of lift of slightly deformable bubbles\cite{ErvinANDTryggvason1997,legendre1998lift}.  We reiterate, though, that for bubble deformation and breakup in shear flows, only a few articles exist: \cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}.  These previous studies mainly examined the dynamics (e.g., rotation angle) of bubble deformation in shear flow.  Concerning bubble breakup, Wei et al.~\cite{WeiQiaXu12} presented one numerical result for a bubble breakup process under the condition of $Ca$ (capillary number) = 35. Sharifi et al\cite{AAMM-16-2} presented two results for bubble breakup corresponding to $Ca=7.5$ and $Ca=11.2$.  We point out that all of the previous computational research on bubble deformation (and breakup) under shear driven flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2} use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (e.g. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}, maintains the gas/liquid interface as sharp, enables accurate approximation of the surface tension force,  and by construction the CLSVOF method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer we must make our computational domain (bigger domain aspect ratio!) thereby adversely effecting the condition number even more for carrying our the pressure projection. 

@@

To highlight the mechanisms of bubble deformation and breakup in a shear flow, we juxtapose the bubble results with those of a drop.  We remark that while the study of critical tensile strength parameters for the bubble is sparse, there have been many studies for the simpler drop problem.  This is because the density ratio for the drop deformation case is almost one so that it is not required that the continuous phase liquid be highly viscous in order to mitigate buoyancy effects.  For completeness, we give a brief overview of previous ``tensile strength'' studies pertaining to drops.

@@

{ Thus, a lot of studies about the deformation and breakup of a drop in simple linear shear have been presented so far. In contrast, few studies have been conducted on bubble deformation and breakup.  We reiterate why there have been few studies regarding the ``tensile strength'' of bubbles. Experimentally, if one wants to isolate the interplay of shearing force with the bubble surface tension force, in the moderate to high Reynolds number regime, and low Capillary number regime, one is restricted to microgravity conditions.  Computational experiments are difficult too. In order to accurately compute the tensile strength of a bubble, one must resort to a combination of parallel computing, the multigrid preconditioned conjugate gradient method\cite{tatebe1993multigrid,SusSmiHusOhtZhi07} for poorly conditioned large sparse matrix systems, adaptive mesh refinement\cite{AMReX_JOSS,SusSmiHusOhtZhi07}, and a robust, volume preserving interface tracking method (we use the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07}).  These aspects of our algorithm allow one to accurately simulate the balance of the restoring bubble surface tension force against the driving wall shear force.
